---
title: "Sherryx_FinalHomeworkCode_05"
author: "Sherry Xie"
date: "2025-04-14"
output: html_document
---

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#loading ggplot for the figures
library(ggplot2)
library(tidyverse)
library(curl)


#loading prettydoc for the theme
#library(prettydoc)
```

## [1] Using the “KamilarAndCooperData.csv” dataset, run a linear regression looking at log(HomeRange_km2) in relation to log(Body_mass_female_mean) and report your 𝛽 coeffiecients (slope and intercept).

```{r}
#Load data
kamilar_cooper <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/refs/heads/master/AN588_Spring25/KamilarAndCooperData.csv")

#fill in blank as N/A
kamilar_cooper <- read.csv(kamilar_cooper, header = TRUE, sep = ",", stringsAsFactors = FALSE)

#show what the data looks like
head(kamilar_cooper)
```

```{r}
#Create new columns in the data set: log(HomeRange) and log(Body Mass)
kamilar_cooper <- kamilar_cooper %>%
  mutate(
    log_HR = log(HomeRange_km2),
    log_BM = log(Body_mass_female_mean)
  ) 

# Fit linear model: log(HomeRange) ~ log(Body Mass)
lm_logHR_logBM <- lm(log_HR ~ log_BM, data = kamilar_cooper)

# Show regression summary
summary(lm_logHR_logBM)

# Show specifically the beta values of intercept and slope
coef(lm_logHR_logBM)
```

## [2] Then, use bootstrapping to sample from your data 1000 times with replacement, each time fitting the same model and calculating the same coefficients. This generates a sampling distribution for each 𝛽 coefficient.

### Estimate the standard error for each of your 𝛽 coefficients as the standard deviation of the sampling distribution from your bootstrap and determine the 95% CI for each of your 𝛽 coefficients based on the appropriate quantiles from your sampling distribution.

```{r}
# Set seed for reproducibility
set.seed(2025)

# Number of bootstrap replicates stated in the instructions 
n_boot <- 1000

# Prepare storage for coefficients
boot_coefs <- matrix(NA, nrow = n_boot, ncol = 2)

# Bootstrap loop
for (i in 1:n_boot) {
  sample_data <- kamilar_cooper[sample(1:nrow(kamilar_cooper), replace = TRUE), ]
  model <- lm(log_HR ~ log_BM, data = sample_data)
  boot_coefs[i, ] <- coef(model)
}

# Convert to dataframe so it will recoginze and run it 
boot_coefs_df <- as.data.frame(boot_coefs)
colnames(boot_coefs_df) <- c("Intercept", "Slope")
```

### How does the former compare to the SE estimated from your entire dataset using the formula for standard error implemented in `lm()`?

```{r}
# Estimate bootstrap standard errors
boot_se <- apply(boot_coefs_df, 2, sd)
boot_se
```

### How does the latter compare to the 95% CI estimated from your entire dataset?

```{r}
# Estimate 95% confidence intervals (2.5% and 97.5% quantiles)
boot_ci <- apply(boot_coefs_df, 2, quantile, probs = c(0.025, 0.975))

# Show results
cat("Bootstrap Standard Errors:\n")
print(round(boot_se, 4))

cat("\nBootstrap 95% Confidence Intervals:\n")
print(round(boot_ci, 4))

```

## EXTRA CREDIT: Write a FUNCTION that takes as its arguments a dataframe, “d”, a linear model, “m” (as a character string, e.g., “logHR\~logBM”), a user-defined confidence interval level, “conf.level” (with default = 0.95), and a number of bootstrap replicates, “n” (with default = 1000). Your function should return a dataframe that includes: beta coefficient names; beta coefficients, standard errors, and upper and lower CI limits for the linear model based on your entire dataset; and mean beta coefficient estimates, SEs, and CI limits for those coefficients based on your bootstrap.

```{r}
# Extra Credit Function
bootstrap_lm_summary <- function(d, m, conf.level = 0.95, n = 1000) {
  # Convert model string to formula
  model_formula <- as.formula(m)
  
  # Fit model to full dataset
  full_model <- lm(model_formula, data = d)
  coef_names <- names(coef(full_model))
  
  # Extract model-based estimates, SEs, and CIs
  full_estimates <- coef(full_model)
  full_se <- summary(full_model)$coefficients[, "Std. Error"]
  full_ci <- confint(full_model, level = conf.level)
  
  # Bootstrapping
  set.seed(2025)
  boot_coefs <- replicate(n, {
    sample_rows <- d[sample(nrow(d), replace = TRUE), ]
    coef(lm(model_formula, data = sample_rows))
  })
  
  boot_coefs <- t(boot_coefs)
  boot_mean <- apply(boot_coefs, 2, mean)
  boot_se <- apply(boot_coefs, 2, sd)
  boot_ci <- apply(boot_coefs, 2, quantile, probs = c((1 - conf.level)/2, 1 - (1 - conf.level)/2))
  
  # Combine everything into a summary dataframe
  result_df <- data.frame(
    Coefficient = coef_names,
    FullModel_Estimate = round(full_estimates, 4),
    FullModel_SE = round(full_se, 4),
    FullModel_CI_Lower = round(full_ci[, 1], 4),
    FullModel_CI_Upper = round(full_ci[, 2], 4),
    Boot_Estimate = round(boot_mean, 4),
    Boot_SE = round(boot_se, 4),
    Boot_CI_Lower = round(boot_ci[1, ], 4),
    Boot_CI_Upper = round(boot_ci[2, ], 4)
  )
  
  return(result_df)
}

```

## EXTRA EXTRA CREDIT: Graph each beta value from the linear model and its corresponding mean value, lower CI and upper CI from a bootstrap as a function of number of bootstraps from 10 to 200 by 10s. HINT: the beta value from the linear model will be the same for all bootstraps and the mean beta value may not differ that much!

```{r}
# Create vector of bootstrap sizes from 10 to 200 by 10
boot_sizes <- seq(10, 200, by = 10)

# Empty dataframe to store results
bootstrap_trace <- data.frame()

# Loop through each bootstrap size
for (n in boot_sizes) {
  set.seed(2025)  # Reproducibility
  boot_coefs <- replicate(n, {
    sample_data <- kamilar_cooper[sample(nrow(kamilar_cooper), replace = TRUE), ]
    coef(lm(log_HR ~ log_BM, data = sample_data))
  })
  
  # Transpose to match structure
  boot_coefs <- t(boot_coefs)
  
  # Extract slope estimates
  slope_vals <- boot_coefs[, "log_BM"]
  
  # Save mean and CI
  bootstrap_trace <- rbind(bootstrap_trace, data.frame(
    n_boot = n,
    mean_slope = mean(slope_vals),
    ci_lower = quantile(slope_vals, 0.025),
    ci_upper = quantile(slope_vals, 0.975)
  ))
}

# Original slope from the full model
original_slope <- coef(lm_logHR_logBM)["log_BM"]

# Plotting the results
library(ggplot2)

ggplot(bootstrap_trace, aes(x = n_boot)) +
  geom_line(aes(y = mean_slope), color = "blue", size = 1.2) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), fill = "skyblue", alpha = 0.3) +
  geom_hline(yintercept = original_slope, linetype = "dashed", color = "red") +
  labs(
    title = "Stability of Bootstrap Estimates for Slope (log_BM)",
    subtitle = "With Increasing Number of Bootstrap Samples",
    x = "Number of Bootstraps",
    y = "Estimated Slope"
  ) +
  theme_minimal()
```

## 5 Struggles I Have Encountered Doing HW 5

1.  extra credit
2.  extra extra credit
